Parsing Postfix log files

### 2007-06-12 doc/logparser.tex
Wrote the background section on Postfix restrictions.

XXX TO BE WRITTEN

>>>>

Postfix restrictions are documented fully in~\cite{smtpd_access_readme,
smtpd_per_user_control, policy-servers}, the following is a brief overview.

Postfix uses several restriction lists (each containing zero or more
restrictions) to decide the fate of an SMTP command, one list for each
stage of the SMTP conversation: client connection, HELO command, MAIL FROM
command, RCPT TO command (possibly multiple times), DATA command, and end
of data.  By default restriction lists will not be evaluated until the
first RCPT TO command is received, because some clients don't deal with
earlier rejections properly; a side benefit of this delay is that Postfix
has more information available when logging the rejection.\footnote{The
\texttt{smtpd\_delay\_reject} parameter controls this behaviour.}  Postfix
uses simple lookup tables as the deciding factor for some restrictions,
e.g. \texttt{check\_client\_access~\nolinebreak
cidr:/etc/postfix/client\_access}\footnote{\texttt{cidr} is the type of the
lookup table.} will check whether the IP address of the connected client
matches the left hand side of each line in the file and will return the
right hand side of the first matching line.  Other restrictions determine
their result in different ways, e.g. \texttt{reject\_rbl\_client
some.rbl.com} checks whether the IP address of the client is present in the
listed RBL.

Each restriction is evaluated to produce a result of \textit{reject\/},
\textit{permit\/}, \textit{dunno\/} or the name of another restriction to
be evaluated.\footnote{Other results are possible, for full details
see~\cite{smtpd_access_readme, smtpd_per_user_control, policy-servers}.}
The meaning of \textit{permit\/} and \textit{reject\/} is fairly obvious;
\textit{dunno\/} means to stop evaluating the current restriction and
continue with the next restriction in the list, allowing more specific
cases to be handled differently from more general cases.  When the result
is another restriction Postfix will evaluate the new restriction, allowing
restrictions to be chosen based on the client IP address, HELO hostname,
sending address, recipient address, etc.  The administrator can define new
restrictions as a list of existing restrictions, allowing arbitrarily long
and complex sequences of lookups and restrictions.

This description is necessarily brief, for further details
see~\cite{smtpd_access_readme, smtpd_per_user_control, policy-servers}.
